General purpose gene annotation using deep learning

### 2020-01-22 paper/main.tex
stuff

\begin{methods}
\section{Methods}
We worked with 192 animal and 60 plant genomes. The data of each genome consists of the latest publically available genomic assembly in form of a FASTA file and the latest annotation in the GFF format. We used all animals genomes in EnsemblMetazoa 45 (?) \citealp{Howe19} as well as those of all non-embargoed plant genomes of the JGI Phytosome 13 database (?, citation). Both data groups were used seperately during training as well as during evaluation. At the start of our work we split both genome groups into training and test sets, which then contained 19 or 6 genomes, respectively. Some of the remaining genomes were used for training purposes and all were used for evaluating our models.

@@

The classes are not balanced equally and the tradeoff between the number of intergenic and intronic bases is quite different in both groups of genomes. Table \ref{Tab:statistics} shows the ratio of class occurences, together with other overall metrics.

@@

During data generation, we only queried for the longest transcript of each gene and did not include FASTA sequences that had no functional annotation (explanation why this is a sign of error?). As Geenuff checks the genomic annotations for potential errors like missing UTR sections and marks areas as potentially erroneous, we were able to effectively blank out those bases during training. For the training, we divided each continous genomic sequence into 20.000 bases long subsequences, which are then used as input or label for the neural network. We appended zero padding if the subsequences are not long enough. If a subsequence is fully marked as erroneous, we do not include it in the training or evaluation datasets.

>>>>

\begin{methods}
\section{Methods}
We worked with 192 animal and 60 plant genomes. The data of each genome consists of the latest publically available genomic assembly in form of a FASTA file and the latest annotation in the GFF format. We used all animals genomes in EnsemblMetazoa 45 (?) \citealp{Howe19} as well as those of all non-embargoed plant genomes of the JGI Phytosome 13 database (?, citation). Both data groups were used seperately during training as well as during evaluation. At the start of our work we split both genome groups into training and test sets. The test sets contained 19 or 6 genomes, respectively. The test genomes were chosen based on being frequently used model species while also representing a broad phylogenetic spread. Some of the remaining genomes were used for training purposes and all were used for evaluating our models.

@@

The classes are not balanced equally and the tradeoff between the number of intergenic and intronic bases is quite different in both groups of genomes. Table \ref{Tab:statistics_all} shows the ratio of class occurences, together with other overall metrics.

@@

	During data generation, we only queried for the longest transcript of each gene and did not include FASTA sequences that had no functional annotation (explanation why this is a sign of error?). As Geenuff checks the genomic annotations for potential errors and is able to mark those areas, we were able to effectively blank out those bases during training. The vast majority of bases masked lie in the intergenic region as the most prevelant error is a missing UTR and Geenuff marks a potentially large intergenic region for it. Table \ref{Tab:statistics_all} and \ref{Tab:statistics_training} show statistics about the masking. For the training, we divided each continous genomic sequence into 20.000 bases long subsequences, which are then used as input or label for the neural network. We appended zero padding if the subsequences are not long enough. If a subsequence is fully marked as erroneous, we do not include it in the training or evaluation datasets.
